When does vapor pressure deficit drive or reduce evapotranspiration?

### 2017-10-04 doc/paper/vpd_et_paper.tex
I took another look at the scatters of LAI vs Delta, VPD, g_a. I think an independent approximation might be okay on for LAI,e ven though at site level there are some correlations (for a few of the sites).
think it's looking realistic to be done with this today

While the relative variability of Term 1 is similiar across PFT, the absolute value of Term 1 varies strongly across PFT. Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT. As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger. Systematic differences in observed temperatures also cause difference in the average magnitude of Term 1. For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term. Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity. To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.

@@

I would say add a table of det dd bar, also evaluated at mean values of each pft, also normalized by std of vpd, and finally

Now time for full plot as a function of t and rh, then break up into single scaling (g, T), and pft-specific plots of (T, RH), or try 3d maps?

\begin{table}
\caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
\centering
\begin{tabular}{l c c c c c}
  \hline
PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)$ & $\frac{\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , D}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
  \hline
CRO & 0.000853 & 0.026241 & 18.523659 & 0.203022 & 0.473311\\
CSH & -0.108234 & -0.091526 & 50.861613 & 0.439379 & 0.931660\\
DBF & -0.012727 & 0.013794 & 19.734435 & 0.164241 & 0.461674\\
ENF & -0.034087 & 0.000706 & 16.611852 & 0.148548 & 0.534425\\
GRA & -0.019637 & -0.000921 & 16.798083 & 0.173552 & 0.631735\\
\hline
\multicolumn{2}{l}{$^{a}$Footnote text here.}

@@

>>>>

While the relative variability of Term 1 is similiar across PFT, the absolute value of Term 1 varies strongly across PFT. Figure \ref{scale_vary}B shows Term 1 evaluated with the mean $g_a$ for each PFT, and at the range of observed temperatures for each PFT. As expected, for the tree PFTs (DBF, ENF) Term 1 is much larger and the temperature dependence is much stronger. Systematic differences in observed temperatures also cause differences in the average magnitude of Term 1. For example, ENF experiences on average colder temperatures and is thus more likely to have a larger scaling term. Additionally, because std$(g_a) \propto \overline{g_a}$, the spread of Term 1 due to $g_a$ variability will be larger, although this is not shown for simplicity. To summarize, the variability of Term 1 will look like Figure \ref{scale_vary}A for each PFT, but the scale of the x and y-axis will increase or decrease according to Figure \ref{scale_vary}B.

@@

Table 3 confirms these expectations for PFT behaviour of $\frac{\partial \; ET}{\partial \; D}$. For all PFTs except for CRO, average $\frac{\partial \; ET}{\partial \; D}$ is less than zero. However, $\frac{\partial \; ET}{\partial \; D}$ evaluated at the average of all variables (e.g. LAI, $T$, $c_s$, $D$) is only negative for CSH and GRA. And, DBF in addition to CRO experiences $\frac{\partial \; ET}{\partial \; D}$ < 0 less than half the time. These observations highlight the effect of the nonlinear function in Figure \ref{term3}: $\frac{\partial \; ET}{\partial \; D}$ has a much steeper slope when the function is negative, and is thus more likely to be large magnitude.

The units of $\frac{\partial \; ET}{\partial \; D}$ make it difficult to interpret if $D$ is even a meaningful contributor to ET's variability. To understand $D$'s contribution better, we use a linear approximation and present $\frac{\partial \; ET}{\partial \; D}$ multiplied by $D$'s standard deviation. The range of $D$'s contribution to ET's variability ranges between 16 - 20 W m$^{-2}$ for all PFTs except for CSH, which is about 51 W m$^{-2}$. Another meaningful comparison is to $\frac{\partial \; ET}{\partial \; R} * std(R)$, as net radiation generally the driver of ET (cite joe berry here). For all PFTs except for CSH $D$ contributes between 14.5 - 20.5 \% of $R$'s contribution to variability. For CSH the portion is much larger, about 44 \%. However it is important to note that a linear approximation about a mean base state is probably not a very good approximation across the range of variabilility, so these values are just estimates of $D$'s contribution to ET's variability.

So far, idealized plots and statistics have illuminated the form of $\frac{\partial \; ET}{\partial \; D}$ and how it varies with PFT. Large mean LAI and uWUE shifts CRO and DBF towards positive $\frac{\partial \; ET}{\partial \; D}$. However, the strongly nonlinear funtion of $\frac{\partial \; ET}{\partial \; D}$ at $\frac{\partial \; ET}{\partial \; D} < 0$ pushes $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative for DBF (it does not do this for CRO because of CRO's high $g1$). ENF's low $g1$ value increases the dependence of $\frac{\partial \; ET}{\partial \; D}$ on $D$, and makes the function more strongly nonlinear. This has the side effect of pushing $\overline{\frac{\partial \; ET}{\partial \; D}}$ negative futher than other PFTs for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$. GRA shows the opposite behavior; a relatively high $g1$  makes the function more linear, decreasing the magnitude of $\overline{\frac{\partial \; ET}{\partial \; D}}$ for a given fraction $\frac{\partial \; ET}{\partial \; D} < 0$ and magntiude $\frac{\partial \; ET}{\partial \; D}(\overline{T,\ldots,D})$ (although $g_a$ and Term 1 also probably have a role in this). Finally, low $uWUE$ of CSH pushes to toward by far the lowest values $\frac{\partial \; ET}{\partial \; D}$ (Figure \ref{term3}). Variability in $D$ accounts for the largest about of $ET$ variability for CSH. For the other PFTs, $D$ contributes less to $ET$ variaibiliyt, but still represents about 15-20 \% of $R$'s contributions to ET variability.

\Begin{table}
\caption{Statistics of $\frac{\partial \; ET}{\partial \; D}$ as a function of PFT.}
\centering
\begin{tabular}{l c c c c c}
  \hline
PFT & $\overline{\frac{\partial \; ET}{\partial \; VPD}}$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)$ & $\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)$ & $\frac{\frac{\partial \; ET}{\partial \; D}\left(\overline{T, \ldots , D}\right)*\text{std}(D)}{ \frac{\partial \; ET}{\partial \; R}\left(\overline{T, \ldots , D}\right)*\text{std}(R)}$ & fraction $\frac{\partial \; ET}{\partial \; VPD} < 0.$ \\
  \hline
CRO & 0.000853 & 0.026241 & 18.523659 & 0.203022 & 0.473311\\
CSH & -0.108234 & -0.091526 & 50.861613 & 0.439379 & 0.931660\\
DBF & -0.012727 & 0.013794 & 19.734435 & 0.164241 & 0.461674\\
ENF & -0.034087 & 0.000706 & 16.611852 & 0.148548 & 0.534425\\
GRA & -0.019637 & -0.000921 & 16.798083 & 0.173552 & 0.631735\\
\hline
\multicolumn{2}{l}{$^{a}$Footnote text here.}

@@

\subsection{Direct observations of $\frac{\partial \; ET}{\partial \; D}$}

Now that we have an intuitive understanding of $\frac{\partial \; ET}{\partial \; D}$'s behavior, we are equipped to intepret fully realistic plots of $\frac{\partial \; ET}{\partial \; D}$ for each PFT. Figure \ref{fig} presents calculated $\frac{\partial \; ET}{\partial \; D}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary.
